---
title: "R Notebook"
output: html_notebook
---
# Weighted Impact Categories per Scenario
Weight the impact categories for a given dataset and scenario.

```{r message=FALSE, warning=FALSE, error=FALSE}
library(readxl)
library(tidyverse)
library(stringr)
library(ggthemes)
library(flextable)


# Specify the impact assessment methodology you're using
methodology_used <- "Environmental Footprint"

# 1. Generalized data processor
process_data <- function(file_path, sheet_name) {
  df <- read_excel(file_path, sheet = sheet_name)
  
  df <- df %>%
    mutate(
      Scenario = factor(Scenario, levels = unique(Scenario)),
      Indicator = factor(Indicator, levels = unique(Indicator)),
      Unit = factor(Unit),
      Quantity = as.numeric(Quantity)
    )
  
  return(df)
}

# 2. Generalized function to calculate weighted categories
calculate_weighted_categories <- function(df, mix_scenario, weighted_tag = "\\(W\\)") {
  df %>%
    filter(
      Scenario == mix_scenario,
      grepl(weighted_tag, Indicator)
    ) %>%
    mutate(
      Indicator = gsub(paste0("\\s*", weighted_tag), "", Indicator),
      Quantity = as.numeric(Quantity)
    ) %>%
    group_by(Indicator) %>%
    summarise(
      total_quantity = sum(Quantity, na.rm = TRUE),
      total_quantity_abs = abs(total_quantity),
      .groups = "drop"
    ) %>%
    arrange(desc(total_quantity_abs)) %>%
    mutate(
      cum_sum = cumsum(total_quantity_abs),
      cum_per = cum_sum / sum(total_quantity_abs),
      first_above80 = which(cum_per >= 0.8)[1],
      above80 = ifelse(row_number() <= first_above80, "yes", "no")
    )
}

create_hotspot_plot <- function(data, add_title = FALSE, title_scenario = NULL) {
  # Get min and max bar values
  min_val <- min(data$total_quantity, na.rm = TRUE)
  max_val <- max(data$total_quantity, na.rm = TRUE)

  # Pad just beyond bars
  lower_limit <- min_val - 0.05 * abs(min_val)
  upper_limit <- max_val + 0.05 * abs(max_val)

  # Use pretty breaks within those limits
  breaks <- pretty(c(lower_limit, upper_limit), n = 6)

  # Build plot
  p <- ggplot(data, aes(x = reorder(Indicator, -total_quantity), y = total_quantity, fill = above80)) +
    geom_bar(stat = "identity") +
    scale_fill_manual(values = c("yes" = "blue", "no" = "lightgrey")) +
    labs(
      x = "Impact Category",
      y = "Weighted Impact (Points)",
      fill = "Hotspot"
    ) +
    theme_few() +
    scale_y_continuous(
      limits = c(min(breaks), max(breaks)),
      breaks = breaks,
      expand = c(0, 0)
    ) +
    coord_flip() +
    theme(
      text = element_text(size = 11),
      axis.title = element_text(size = 11),
      axis.text = element_text(size = 11)
    )

  if (add_title && !is.null(title_scenario)) {
    p <- p + ggtitle(paste("Most important impact categories\n in", title_scenario))
  }

  return(p)
}


```

# Test


```{r message=FALSE, warning=FALSE, error=FALSE}
# test 

df <- process_data("EF_2028f.xlsx","LCIA RawData")
# calculate_weighted_categories <- function(df, mix_scenario, weighted_tag = "\\(W\\)")
weighted_impacts <- calculate_weighted_categories(df, "100% grid, 2028", weighted_tag = "\\(W\\)")

create_hotspot_plot(data = weighted_impacts, add_title = TRUE,title_scenario ="100% grid, 2028")
```

Run the code over all scenarios to get graphs for each one. Save each in Figures folder.

```{r}

library(purrr)
library(tidyverse)

# Get unique scenarios
scenarios <- unique(df$Scenario)

# Run analysis and plotting for each scenario
results <- map(scenarios, function(scenario) {
  # Calculate weighted categories
  weighted_data <- calculate_weighted_categories(df, scenario)
  
  # Create plot
  plot <- create_hotspot_plot(weighted_data, add_title = TRUE, title_scenario = scenario)
  
  # Save plot (sanitize file name)
  safe_name <- gsub("[^a-zA-Z0-9]", "_", scenario)
  ggsave(paste0("Figures/hotspot_", safe_name, ".png"), plot = plot, bg = "white")
  
  # Return both data and plot if needed
  list(
    scenario = scenario,
    data = weighted_data,
    plot = plot
  )
})

```
# Hotspot life cycle phases

```{r}
library(readxl)
library(tidyverse)

# -------------------------
# Generalized LCP Hotspot Function
# -------------------------
plot_lcp_stacked_all_scenarios <- function(file_path,
                                           sheet_name = "LCIA RawData",
                                           indicator,
                                           output_path = NULL) {

  if (is.null(output_path)) {
    safe_indicator <- gsub("[^a-zA-Z0-9]", "_", indicator)
    output_path <- paste0("Figures/stacked_all_LCPs_", safe_indicator, ".png")
  }

  # Read and process data
  raw_data <- read_excel(file_path, sheet = sheet_name)

  df <- raw_data %>%
    mutate(
      Scenario = factor(Scenario),
      Indicator = factor(Indicator),
      Unit = factor(Unit),
      Quantity = as.numeric(Quantity)
    )

  # Filter and prepare
  burdens_df <- df %>%
    filter(Indicator == indicator, LifeCyclePhase != "Credits")

  # Get data for stacked bars
  summary_df <- burdens_df %>%
    group_by(Scenario, LifeCyclePhase) %>%
    summarise(climate_change = sum(Quantity, na.rm = TRUE), .groups = "drop")

  # Calculate 80% thresholds per scenario
  thresholds <- burdens_df %>%
    group_by(Scenario) %>%
    summarise(
      total_burden = sum(Quantity, na.rm = TRUE),
      eighty_percent = 0.8 * total_burden,
      .groups = "drop"
    )

  # Merge x positions for geom_segment (ensure matching factor levels)
  scenario_levels <- levels(summary_df$Scenario)

  thresholds <- thresholds %>%
    mutate(
      x = as.numeric(factor(Scenario, levels = scenario_levels)) - 0.4,  # left edge
      xend = as.numeric(factor(Scenario, levels = scenario_levels)) + 0.4  # right edge
    )

  # Plot
  plot <- ggplot(summary_df, aes(x = Scenario, y = climate_change, fill = LifeCyclePhase)) +
    geom_col(position = "stack") +
    geom_segment(data = thresholds,
                 aes(x = x, xend = xend, y = eighty_percent, yend = eighty_percent),
                 inherit.aes = FALSE,
                 linetype = "dashed",
                 color = "black") +
    geom_text(data = thresholds,
              aes(x = as.numeric(factor(Scenario)), y = eighty_percent, label = "80%"),
              inherit.aes = FALSE,
              vjust = -0.5, size = 3, color = "black") +
    labs(
      x = "Scenario",
      y = paste(indicator, "(EF 3.1)"),
      fill = "Life Cycle Phase"
    ) +
    theme_minimal(base_size = 11) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "bottom",
      legend.text = element_text(size = 11)
    ) +
    guides(fill = guide_legend(nrow = 2, byrow = TRUE)) +
    scale_fill_brewer(
      palette = "Spectral",
      labels = c("Manufacture", "Disposal", "Distribution", "Resource Extraction", "Use")
    )

  # Save
  ggsave(output_path, plot = plot, bg = "white", width = 10, height = 6)

  return(plot)
}



```

## Test
The most relevant life cycle stages in each scenario are those that contribute cumulatively more than 80% to any impact category (before normalization and weighting).

```{r}
plot_lcp_stacked_all_scenarios(
  file_path = "EF_2028f.xlsx",
  indicator = "Climate change - Total"
)
```






```{r}
# Step 1: Read full data
df <- process_data("EF_2028f.xlsx", "LCIA RawData")
filter_allowed_indicators <- function(df, methodology_used) {
  # Define valid EF indicators
  allowed_ef_indicators <- c(
    "Climate change - Total",
    "Carcinogenic effects - Total",
    "Non-carcinogenic effects - Total",
    "Respiratory effects, inorganics",
    "Ozone layer depletion",
    "Photochemical ozone creation",
    "Ionizing radiation",
    "Freshwater ecotoxicity - Total",
    "Freshwater eutrophication",
    "Marine eutrophiation",
    "Terrestrial eutrophiation",
    "Freshwater and terrestrial acidification",
    "Fossils",
    "Minerals and metals",
    "Land use",
    "Water scarcity"
  )

  # If using EF, apply the filter
  if (methodology_used == "Environmental Footprint") {
    df <- df %>% filter(Indicator %in% allowed_ef_indicators)
  }

  # For other methods, return df unchanged
  return(df)
}
# Step 1: Read the raw dataset
df <- process_data("EF_2028f.xlsx", "LCIA RawData")

# Step 2: Apply filtering logic specific to the selected methodology
relevance_df <- filter_allowed_indicators(df, methodology_used)

# Step 3: Run relevance analysis
indicators <- unique(relevance_df$Indicator)

relevant_all_indicators <- purrr::map_dfr(indicators, function(ind) {
  get_relevant_life_cycle_stages(relevance_df, ind)
}) %>%
  arrange(Scenario, Indicator)

ft <- relevant_all_indicators %>%
  flextable() %>%
  set_header_labels(
    Scenario = "Scenario",
    Indicator = "Indicator",
    LifeCyclePhase = "Life Cycle Phase",
    total_quantity = "Impact",
    cum_per = "Cumulative %"
  ) %>%
  # Format Impact as scientific with 2 decimal places
 flextable::colformat_scientific(j = "total_quantity", digits = 2) %>%
  # Format Cumulative % as normal decimal with 2 decimal places
  colformat_num(j = "cum_per", digits = 2) %>%
  autofit() %>%
  theme_vanilla() %>%
  set_caption("Most Relevant Life Cycle Phases (≥80%) per Indicator and Scenario")

```



# Listing out the most relevant life cycle stages in each scenario for a given impact category.

```{r}
library(dplyr)

get_relevant_life_cycle_stages <- function(df, indicator_name) {
  df %>%
    filter(Indicator == indicator_name, LifeCyclePhase != "Credits", !is.na(Quantity)) %>%
    group_by(Scenario, LifeCyclePhase) %>%
    summarise(total_quantity = sum(Quantity, na.rm = TRUE), .groups = "drop") %>%
    group_by(Scenario) %>%
    arrange(desc(abs(total_quantity))) %>%
    mutate(
      total_abs = sum(abs(total_quantity), na.rm = TRUE),
      cum_sum = cumsum(abs(total_quantity)),
      cum_per = cum_sum / total_abs,
      row_num = row_number(),
      cutoff = which(cum_per >= 0.8)[1],
      relevant = row_num <= cutoff
    ) %>%
    filter(relevant) %>%
    mutate(Indicator = indicator_name) %>%
    select(Scenario, Indicator, LifeCyclePhase, total_quantity, cum_per)
}



# Example usage

df <- process_data("EF_2028f.xlsx", sheet_name = "LCIA RawData")
# Get list of all unique indicators
indicators <- unique(df$Indicator)

# Loop over each and bind results into a single table
relevant_all_indicators <- purrr::map_dfr(indicators, function(ind) {
  get_relevant_life_cycle_stages(df, ind)
})

ggplot(relevant_all_indicators, aes(x = Scenario, y = LifeCyclePhase)) +
  geom_tile(aes(fill = Indicator)) +
  facet_wrap(~ Indicator, scales = "free", ncol = 2) +
  labs(
    title = "Relevant Life Cycle Phases (≥80%)",
    x = "Scenario",
    y = "Life Cycle Phase"
  ) +
  theme_minimal(base_size = 11) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```




